noam_data_exploration_v1

### 2024-08-22 noam_data_exploration_v3.Rmd
Filtered model effects for diversity-stability model to plot only over the ranges of the data

@@

richness_cv_mod <- glmmTMB(cover_cv ~ richness_mean*habitat*island_side + (1|site/transect), family = Gamma(link = "log"),
                          data = diversity_stability_quadrat, na.action = na.pass)
# model.sel(dredge(richness_cv_mod)) # top model is as specified, second best (delta = 10) doesn't include rich:hab:coast
r.squaredGLMM(richness_cv_mod) # 0.61, 0.64
summary(richness_cv_mod)
emmeans(richness_cv_mod, pairwise ~ island_side | habitat, var = "richness_mean")
richness_cv_emm <- emmip(richness_cv_mod,
                        habitat ~ richness_mean | island_side,
                        at = list(richness_mean = seq(0,5,0.1)), plotit = F, CIs = T)

(richness_cv_plot = 
  ggplot() +
  geom_point(data = diversity_stability_quadrat,
             aes(x = richness_mean, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = richness_cv_emm, aes(x = richness_mean, y = exp(yvar), colour = island_side), size = 2) +
  geom_ribbon(data = richness_cv_emm, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = island_side), alpha = 0.3) +
#   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +
  scale_y_log10() +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

(shannon_cv_plot_2 = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# lter_2 obscures a lot of the trends

@@

(loreau_stability_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = loreau_synchrony, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = synchrony_cv_emm, aes(x = loreau_synchrony, y = exp(yvar), colour = island_side), size = 1) +
  geom_ribbon(data = synchrony_cv_emm, aes(x = loreau_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = island_side),
             alpha = 0.3) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +
   scale_y_log10() +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

>>>>

## FOR FILTERING MODEL EFFECTS LATER
diversity_ranges <- diversity_stability_quadrat %>%
  group_by(island_side, habitat) %>%
  summarise(min_richness = min(richness_mean),
            max_richness = max(richness_mean),
            min_shannon = min(shannon_mean),
            max_shannon = max(shannon_mean),
            min_loreau = min(loreau_synchrony),
            max_loreau = max(loreau_synchrony))

@@

richness_cv_mod <- glmmTMB(cover_cv ~ richness_mean*habitat*island_side + (1|site/transect), family = Gamma(link = "log"),
                          data = diversity_stability_quadrat, na.action = na.pass)
# model.sel(dredge(richness_cv_mod)) # top model is as specified, second best (delta = 10) doesn't include rich:hab:coast
r.squaredGLMM(richness_cv_mod) # 0.61, 0.64
summary(richness_cv_mod)
emtrends(richness_cv_mod, pairwise ~ island_side | habitat, var = "richness_mean")
richness_cv_emm <- emmip(richness_cv_mod,
                        habitat ~ richness_mean | island_side,
                        at = list(richness_mean = seq(0,5,0.1)), plotit = F, CIs = T)

filtered_richness_effects <- richness_cv_emm %>%
  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)

(richness_cv_plot = 
  ggplot() +
  geom_point(data = diversity_stability_quadrat,
             aes(x = richness_mean, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_richness_effects, aes(x = richness_mean, y = exp(yvar), colour = island_side), size = 2) +
  geom_ribbon(data = filtered_richness_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = island_side), alpha = 0.3) +
#   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +
  scale_y_log10() +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

shannon_cv_mod <- glmmTMB(cover_cv ~ shannon_mean*habitat*island_side + (1|site/transect), family = Gamma(link = "log"),
                          data = diversity_stability_quadrat, na.action = na.pass)
# model.sel(dredge(shannon_cv_mod)) # top model is as specified, second best (delta = 10) doesn't include shann:hab:coast
r.squaredGLMM(shannon_cv_mod) # 0.55, 0.59
summary(shannon_cv_mod)
emtrends(shannon_cv_mod, pairwise ~ island_side | habitat, var = "shannon_mean")
shannon_cv_emm <- emmip(shannon_cv_mod,
                        habitat ~ shannon_mean | island_side,
                        at = list(shannon_mean = seq(0,1.25,0.05)), plotit = F, CIs = T)

filtered_shannon_effects <- shannon_cv_emm %>%
  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
  filter(shannon_mean >= min_shannon & shannon_mean <= max_shannon)

(shannon_cv_plot = 
  ggplot() +
  geom_point(data = diversity_stability_quadrat,
             aes(x = shannon_mean, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_shannon_effects, aes(x = shannon_mean, y = exp(yvar), colour = island_side), size = 2) +
  geom_ribbon(data = filtered_shannon_effects, aes(x = shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = island_side), alpha = 0.3) +
#   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +
  scale_y_log10() +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

filtered_synchrony_effects <- synchrony_cv_emm %>%
  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
  filter(loreau_synchrony >= min_loreau & loreau_synchrony <= max_loreau)

(loreau_stability_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = loreau_synchrony, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_synchrony_effects, aes(x = loreau_synchrony, y = exp(yvar), colour = island_side), size = 1) +
  geom_ribbon(data = filtered_synchrony_effects, aes(x = loreau_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = island_side),
             alpha = 0.3) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +
   scale_y_log10() +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
